Capstone Movielens Report

### 2025-05-07 reports/capstone-movielens-report.draft4.Rmd
Draft report, section fixed: Overall Mean Rating (Naive) Model.

```{r eval=FALSE }
open_logfile <- function(file_name){
  log_file_name <- as.character(Sys.time()) |> 
    str_replace_all(':', '_') |> 
    str_replace(' ', 'T') |>
    str_c(file_name)

@@

print_log <- function(msg){
  print(str_glue(msg))
}
put_log <- function(msg){
  put(str_glue(msg))
}
put_log1 <- function(msg_template, arg1){
  msg <- str_replace_all(msg_template, "%1", as.character(arg1))
  put(str_glue(msg))
}
put_log2 <- function(msg_template, arg1, arg2){
  msg <- msg_template |> 
    str_replace_all("%1", as.character(arg1)) |>
    str_replace_all("%2", as.character(arg2)) |>
    str_glue()

  put(msg)
}

@@

```{r pressure, echo=TRUE, eval=FALSE}
  mu <- mean(edx$rating)
```
```{r}
print(mu)
```

If we predict all unknown ratings with $\hat{\mu}$, we obtain the following RMSE: 
```{r eval=FALSE}
mu.MSEs <- naive_model_MSEs(mu)
```
```{r}
  data.frame(fold_No = 1:5, MSE = mu.MSEs) |>
    data.plot(title = "MSE resuls of the 5-fold CV method applied to the Overall Mean Rating Model",
                xname = "fold_No", 
                yname = "MSE")

@@

```{r eval=FALSE}
naive_model_MSEs <- function(val) {
  sapply(edx_CV, function(cv_item){
    mse(cv_item$validation_set$rating - val)
  })
}
```

One more function we will need for further analysis of the current model is the [`naive_model_RMSE` one](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R#L59) defined in the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R) script already mentioned above:
```{r}
naive_model_RMSE <- function(val){
  sqrt(mean(naive_model_MSEs(val)))
}
```

@@

If we plug in any other number, we will get a higher RMSE. Let's prove that by the following small investigation:
```{r eval=FALSE}
  deviation <- seq(0, 6, 0.1) - 3

  deviation.RMSE <- sapply(deviation, function(delta){
    naive_model_RMSE(mu + delta)
  })
  put_log1("RMSE values have been computed for %1 deviations from the Overall Mean Rating.",
           length(deviation))
```

@@

put_log1("Minimum RMSE is achieved when the deviation from the mean is: %1",
         which_min_deviation)

put_log1("Is the previously computed RMSE the best for the current model: %1",
         mu.RMSE == min_rmse)

>>>>

```{r eval=FALSE }
## Logging Helper functions -----------------------------------------------------
open_logfile <- function(file_name){
  log_file_name <- as.character(Sys.time()) |> 
    str_replace_all(':', '_') |> 
    str_replace(' ', 'T') |>
    str_c(file_name)

@@

msg.set_arg <- function(msg_template, arg, arg.name = "%1") {
  msg_template |> 
    str_replace_all(arg.name, as.character(arg))
}
msg.glue <- function(msg_template, arg, arg.name = "%1"){
  msg_template |>
    msg.set_arg(arg, arg.name) |>
    str_glue()
}

print_log <- function(msg){
  print(str_glue(msg))
}
put_log <- function(msg){
  put(str_glue(msg))
}

get_log1 <- function(msg_template, arg1) {
  str_glue(str_replace_all(msg_template, "%1", as.character(arg1)))
}
print_log1 <- function(msg_template, arg1){
  print(get_log1(msg_template, arg1))
}
put_log1 <- function(msg_template, arg1){
  put(get_log1(msg_template, arg1))
}

get_log2 <- function(msg_template, arg1, arg2) {
  msg_template |> 
    str_replace_all("%1", as.character(arg1)) |>
    str_replace_all("%2", as.character(arg2)) |>
    str_glue()
}
print_log2 <- function(msg_template, arg1, arg2){
  print(get_log1(msg_template, arg1, arg2))
}
put_log2 <- function(msg_template, arg1, arg2){
  put(get_log1(msg_template, arg1, arg2))
}

# ...
```

@@

```{r pressure, echo=TRUE}
mu <- mean(edx$rating)
print(mu)
```

If we predict all unknown ratings with $\hat{\mu}$, we obtain the following RMSE: 
```{r}
mu.MSEs <- naive_model_MSEs(mu)
data.frame(fold_No = 1:5, MSE = mu.MSEs) |>
  data.plot(title = "MSE resuls of the 5-fold CV method applied to the Overall Mean Rating Model",
              xname = "fold_No", 
              yname = "MSE")

@@

```{r eval=FALSE }
naive_model_MSEs <- function(val) {
  sapply(edx_CV, function(cv_item){
    mse(cv_item$validation_set$rating - val)
  })
}
```

One more function we will need for further analysis of the current model is the [`naive_model_RMSE` one](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/914dd47023dc21b6b06a0ca412b7f7281602275a/r/src/support-functions/common-helper.functions.R#L59) defined in the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R) script already mentioned above:
```{r eval=FALSE }
naive_model_RMSE <- function(val){
  sqrt(mean(naive_model_MSEs(val)))
}
```

@@

If we plug in any other number, we will get a higher RMSE. Let's prove that by the following small investigation:
```{r }
  deviation <- seq(0, 6, 0.1) - 3

  deviation.RMSE <- sapply(deviation, function(delta){
    naive_model_RMSE(mu + delta)
  })
```

@@

print_log1("Minimum RMSE is achieved when the deviation from the mean is: %1",
         which_min_deviation)

print_log1("Is the previously computed RMSE the best for the current model: %1",
         mu.RMSE == min_rmse)
